### 2016-10-14 mcmc.tex
swap section order
mcmc

% to-do
% -----
% - Deal with all mentions of ``DFM'' or ``HOGG''
% - Go through github issues, if any.
% - Systematically cite chapters of Brooks \etal in each section.
% - Skim MacKay and make sure we cite him in any relevant place.
% - Mention & cite STAN in methods section
% - Need to mention case where there is an MC integral INSIDE the likelihood function
% - Need to mention testing the likelihood function, by, say, 1-d scans back and forth.
% - Make sure we say ``don't use MCMC for 1-d problems or if you have alternatives to directly integrate!''
% - Would love to have the Note symbols to be links to the notes and the notes to have links back to the main text...!
% - Send to friendlies for comments
%   - Gelman
%   - Goodman
%   - Lang
%   - Rix
%   - Bovy
%   - Brewer
% - Kill all issues in this list, above, and then:
%   - Post on arXiv
%   - Blog about it

@@

\begin{document}\sloppy\sloppypar\raggedbottom\thispagestyle{plain}%
%
\section*{Data analysis recipes:\\ \chaptertitle\footnotemark}
%
\footnotetext{%
    The notes begin on page~\pageref{note:first}, including the
    license\note{\label{note:first}
        Copyright 2013, 2014, 2015, 2016 the authors (OMG this took us a long time).
        This work is licensed under a
        \href{http://creativecommons.org/licenses/by-nc-nd/3.0/deed.en\_US}{%
            Creative Commons Attribution-NonCommercial-NoDerivs 3.0 Unported
            License}.}
    and the acknowledgements\note{%
        We would like to thank
          Jo Bovy (IAS),
          Brendon Brewer (Auckland),
          Will Farr (Birmingham),
          Jonathan Goodman (NYU),
          Fengji Hou (NYU),
          Dustin Lang (CMU),
          Phil Marshall (KIPAC), and
          Iain Murray (Edinburgh)
        for valuable advice and comments over the years, and
          Alessandro Gentilini
        for a very close read.
        This project benefitted from various research grants, especially
          NSF grants AST-0908357, IIS-1124794, and AST-1517237,
          NASA grant NNX12AI50G,
          and the The Moore-Sloan Data Science Environment at NYU.}
}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.5\textwidth]{problems/MH}
\end{center}
\caption{Solution to \problemname~\ref{prob:MH}:
MCMC samples from a Gaussian (black) and the true distribution (blue).}
\label{fig:MH}
\end{figure}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.5\textwidth]{problems/MH2}
\end{center}
\caption{Solution to \problemname~\ref{prob:MH2}:
MCMC samples from a uniform distribution (black) and the true distribution
(blue).}
\label{fig:MH2}
\end{figure}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{problems/twod_a}
\end{center}
\caption{Solution to \problemname~\ref{prob:twod}a:
MCMC samples from a two-dimensional Gaussian (scatter plot) and the
one-dimensional marginalized distributions (histograms).}
\label{fig:twod-a}
\end{figure}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.5\textwidth]{problems/logprior}
\end{center}
\caption{Solution to \problemname~\ref{prob:logprior}:
The black histogram shows the MCMC samples and the blue curve is the analytic
target density.}
\label{fig:logprior}
\end{figure}

@@

It is also worth noting that there is not just one autocorrelation function or
autocorrelation time for an MCMC chain.
Instead, there is an autocorrelation time $\tau_f$ for every function
$f(\pars)$ and these will, in general, be different.
Therefore, the autocorrelation time used in \equationname~(\ref{eq:tauvar}) is
the time computed on the chain of $q(\pars)$ values instead of the
autocorrelation time for $\pars$.

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.5\textwidth]{problems/itertau}
\end{center}
\caption{Solution to \problemname~\ref{prob:itertau}:
Estimates of the integrated autocorrelation time of different segments of the
MCMC chain from \problemname~\ref{prob:MH} (black lines) and for the full
chain (orange line) as a function of window size $M$.
The ``optimized'' value computed using an iterative procedure is overplotted
as a dashed line and its value is listed in the title.}
\label{fig:itertau}
\end{figure}

>>>>

% to-do
% -----
% - Deal with all mentions of ``DFM'' or ``HOGG''
% - Go through github issues, if any.
% - Need to mention case where there is an MC integral INSIDE the likelihood function
% - Need to mention testing the likelihood function, by, say, 1-d scans back and forth.
% - Send to friendlies for comments
%   - Gelman
%   - Goodman
%   - Lang
%   - Rix
%   - Bovy
%   - Brewer
% - Kill all issues in this list, above, and then:
%   - Post on arXiv
%   - Blog about it

@@

\begin{document}\sloppy\sloppypar\raggedbottom\frenchspacing\thispagestyle{plain}%
%
\section*{Data analysis recipes:\\ \chaptertitle\footnotemark}
%
\footnotetext{%
    The notes begin on page~\pageref{note:first}, including the
    license\note{\label{note:first}
        Copyright 2013, 2014, 2015, 2016 the authors (OMG this took us a long time).
        This work is licensed under a
        \href{http://creativecommons.org/licenses/by-nc-nd/3.0/deed.en\_US}{%
            Creative Commons Attribution-NonCommercial-NoDerivs 3.0 Unported
            License}.}
    and the acknowledgements\note{%
        We would like to thank
          Jo Bovy (IAS),
          Brendon Brewer (Auckland),
          Will Farr (Birmingham),
          Jonathan Goodman (NYU),
          Fengji Hou (NYU),
          Dustin Lang (CMU),
          Phil Marshall (KIPAC), and
          Iain Murray (Edinburgh)
        for valuable advice and comments over the years, and
          Alessandro Gentilini
        for a very close read.
        This project benefitted from various research grants, especially
          NSF grants AST-0908357, IIS-1124794, and AST-1517237,
          NASA grant NNX12AI50G,
          and the The Moore-Sloan Data Science Environment at NYU.}
}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.5\textwidth]{problems/MH}
\end{center}
\caption{Solution to \problemname~\ref{prob:MH}:
\MCMC\ samples from a Gaussian (black) and the true distribution (blue).}
\label{fig:MH}
\end{figure}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.5\textwidth]{problems/MH2}
\end{center}
\caption{Solution to \problemname~\ref{prob:MH2}:
\MCMC\ samples from a uniform distribution (black) and the true distribution
(blue).}
\label{fig:MH2}
\end{figure}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{problems/twod_a}
\end{center}
\caption{Solution to \problemname~\ref{prob:twod}a:
\MCMC\ samples from a two-dimensional Gaussian (scatter plot) and the
one-dimensional marginalized distributions (histograms).}
\label{fig:twod-a}
\end{figure}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.5\textwidth]{problems/logprior}
\end{center}
\caption{Solution to \problemname~\ref{prob:logprior}:
The black histogram shows the \MCMC\ samples and the blue curve is the analytic
target density.}
\label{fig:logprior}
\end{figure}

@@

It is also worth noting that there is not just one autocorrelation function or
autocorrelation time for an \MCMC\ chain.
Instead, there is an autocorrelation time $\tau_f$ for every function
$f(\pars)$ and these will, in general, be different.
Therefore, the autocorrelation time used in \equationname~(\ref{eq:tauvar}) is
the time computed on the chain of $q(\pars)$ values instead of the
autocorrelation time for $\pars$.

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.5\textwidth]{problems/itertau}
\end{center}
\caption{Solution to \problemname~\ref{prob:itertau}:
Estimates of the integrated autocorrelation time of different segments of the
\MCMC\ chain from \problemname~\ref{prob:MH} (black lines) and for the full
chain (orange line) as a function of window size $M$.
The ``optimized'' value computed using an iterative procedure is overplotted
as a dashed line and its value is listed in the title.}
\label{fig:itertau}
\end{figure}
